---
title: "Generalized with Marginal"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Model fitting
 
 Here we allow for the user to input an array of people x ages x states with the exact position of an individual in every age, and a matrix of potential transitions, to train the model. The marginal computation over all states is then entry in the matrix product of the j,k position at age N, but you can also use the conditional computations.
 
* load data

```{r}
library(MSGene)
library(data.table)
dfh <- readRDS(system.file("data", "msgene_sampledf.rds", package = "MSGene"))
train=dfh[1:(nrow(dfh)*0.80),]
```

# set transition matrix 
```{r}
nstates=c("Health","Ht","HyperLip","Dm","Cad","death","Ht&HyperLip","HyperLip&Dm","Ht&Dm","Ht&HyperLip&Dm")

transitions=matrix(0,nrow = length(nstates),ncol=length(nstates))
rownames(transitions)=nstates
colnames(transitions)=nstates
transitions["Health","Health"]=1
transitions["Health","Ht"]=1
transitions["Health","HyperLip"]=1
transitions["Health","Dm"]=1
transitions["Health","Cad"]=1
transitions["Health","death"]=1

transitions["Ht","Ht"]=1
transitions["Ht","Ht&HyperLip"]=1
transitions["Ht","Ht&Dm"]=1
transitions["Ht","Cad"]=1
transitions["Ht","death"]=1


transitions["HyperLip","HyperLip"]=1
transitions["HyperLip","Ht&HyperLip"]=1
transitions["HyperLip","HyperLip&Dm"]=1
transitions["HyperLip","Cad"]=1
transitions["HyperLip","death"]=1

transitions["Dm","Dm"]=1
transitions["Dm","Ht&Dm"]=1
transitions["Dm","HyperLip&Dm"]=1
transitions["Dm","Cad"]=1
transitions["Dm","death"]=1

transitions["Ht&HyperLip","Ht&HyperLip"]=1
transitions["Ht&HyperLip","Ht&HyperLip&Dm"]=1
transitions["Ht&HyperLip","Cad"]=1
transitions["Ht&HyperLip","death"]=1

transitions["Ht&Dm","Ht&Dm"]=1
transitions["Ht&Dm","Ht&HyperLip&Dm"]=1
transitions["Ht&Dm","Cad"]=1
transitions["Ht&Dm","death"]=1


transitions["HyperLip&Dm","HyperLip&Dm"]=1
transitions["HyperLip&Dm","Ht&HyperLip&Dm"]=1
transitions["HyperLip&Dm","Cad"]=1
transitions["HyperLip&Dm","death"]=1


transitions["Ht&HyperLip&Dm","Ht&HyperLip&Dm"]=1
transitions["Ht&HyperLip&Dm","Cad"]=1
transitions["Ht&HyperLip&Dm","death"]=1


## absorbing states
transitions["death","death"]=1
transitions["Cad","Cad"]=1

print(transitions)
```

# Indicate your at risk state for the years of interest 


```{r,warning=FALSE,message=FALSE}
ages=c(40:81)
s=statusarray(df_frame = data.table(train),ages = ages,nstates = nstates)
```

```{r}
dim(s)
head(s[,18,])
```

## now fit the model which will expand over the potential transitions in the transition matrix according to your covariates list

```{r,warning=FALSE,message=FALSE}
ages=40:80
covariates="cad.prs+f.31.0.0+smoke+antihtn_now+statin_now"
train$cad.prs=scale(train$cad.prs)
mfit=modelfitfun(ages = ages,transit_mat=transitions,covariates,df = train,statusarray = s)
```

  
## now we will need to extract coefficients for each age and relevant state to state transition 

```{r,warning=FALSE,message=FALSE}
ages=40:80
coeff_array=coef_array_func(ages = ages,transit_mat = transitions,modelfit = mfit)
```

## now we have an array that is ages x start 
```{r}
dim(coeff_array)
```


## then we will need to compute a transition matrix for each age and persno

here is for one person with a particular covariate profile at age 50

```{r}
age=50
covariates=c(1,1,1,0,0,0)
absorbing_states=c(5,6)


age_specific_matrix <- generate_single_transition_matrix(age = 50, coeff_array, covariates, absorbing_states = c(5, 6))
print(age_specific_matrix)

```

Note that the probs for each row sum to 1 and the for the absorbing states the j,j entry is 1.

```{r}
rowSums(age_specific_matrix)
age_specific_matrix[absorbing_states[1],absorbing_states[1]]
age_specific_matrix[absorbing_states[2],absorbing_states[2]]
```

## Now let's compute over an entire interval: 

```{r}
age_start <- 40
age_end <- 80
indcovariates=c(1,1,1,0,0,0)
# Execute the function.


result_matrix <- calculate_matrix_product_over_interval(coeff_array, indcovariates, age_start, age_end,absorbing_states = c(5,6))

# result_list=list()
# for(i in 1:nrow(atrisk)){
#   result_list[[i]]=calculate_matrix_product_over_interval(coeff_array, atrisk[i,], age_start, age_end,absorbing_states = c(5,6))
# }


print(result_matrix)
```

Great!  But can also compute the lifetime risk as a product of one step and compare differences with our earlier 'conditional' method:

```{r}
cmat=coeff_array[,"Health","Cad",]

intercept=1
sex=c(0,1)
cad.prs=1
smoke=0
antihtn_now=0
statin_now=0
atrisk=expand.grid(intercept,cad.prs,sex,smoke,antihtn_now,statin_now) 

mso = compute_prediction_product_matrix(
coefmat = cmat,
atrisk = atrisk,
agepredinterval = c(40:80)
)

result1=mso$PredictedIntervalrisk[2]


## compare

covariates="cad.prs+f.31.0.0+smoke+antihtn_now+statin_now"
modelfit = fitfunc2(
  df_frame=data.table::data.table(train),
  ages = ages,
  nstates = nstates,
  mode = "binomial",
  covariates = covariates)
```

### Look at all that we've produced! 

```{r}
attributes(modelfit)
```

You can see in this array we have the number of events, the absolute rate per year per transition, the number at risk per year and per transition, and the list of models over years and states.

### Extract coefficients for health-->CAD transition

To compute our risk calculations, we need to extract the unsmoothed coefficients, we will feed through a smoother, and then use these coefficients to compute our product. We want to check to make sure these match!

```{r}
a = coefplotsmooth2(
    ages = ages,
    start = nstates[1],
    stop = "Cad",
    modelfit = modelfit,
    window_width = 20,
    span = 0.75,
    degree = 2
  )
healthy_coefs = a$custom_smooth



mso = compute_prediction_product_matrix(
coefmat = healthy_coefs,
atrisk = atrisk,
agepredinterval = c(40:80)
)

result2=mso$PredictedIntervalrisk[2]
all.equal(result1,result2)
```

## 'result_matrix' now contains the product of transition matrices over the specified age interval.

Confirm that result matches the product of individual matrices:

```{r}
covariates=c(1,1,1,0,0,0)
ml=mat.test(coefficients = coeff_array,covariates = covariates,age_start = 40,age_end = 79,absorbing_states = c(5,6))

cumulative_product <- Reduce(function(x, y) x %*% y, ml)

all.equal(cumulative_product,result_matrix)
```
